---
title: "Morality in the time of cognitive famine - Public Goods Game"
author: Jonas Kristoffer Lindeløv
date: December, 2018

always_allow_html: yes

output:
  github_document:
    toc: true

---
<!--
# TO DO
* Number of trials per subject as a criterion?
-->

# About
This is part of the analysis that accompanies the paper "Morality in the time of cognitive famine" by Panos, Jonas, Michaela, and others. You are now looking at the analysis of **experiment 1** using the Public Goods Game to study cooperation and Operation Span task to load working memory.


# Setting up
Load appropriate stuff:
```{r setup, results=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lme4)
#library(kableExtra)

source('misc/functions utility.R')
source('misc/functions inference.R')  # Contains LRT and LRT_binom
```

You could redo the preprocessing of the original data if you wanted to. It saves the .Rda files which are loaded in the sections below.
```{r, eval=FALSE, warning=FALSE, message=FALSE}
source('preprocess PGG.R')
```


# Load data
... and remove two subjects who did not follow instructions (noted by the research assistant)
```{r load data}
# Load the data.frame (not in csv because the matrix columns would be lost)
D = readRDS('data/pgg.Rda') %>%
  filter(condition == 'experiment') %>%  # Remove practice, etc.
  select(-encode, -equationCorrect, -equationScores, -equationRTs, -recallAns, -equationAnss)  # Tidyr doesn't like these matrix columns and we won't be using them
```

# Descriptives
For manuscript:
```{r exp12 descriptives}
D_id = D[!duplicated(D$id), ]

# Per-group descriptives for paper
D %>%
  # One row per participant
  group_by(id) %>%
  slice(1) %>%
  
  # Per-experiment group
  group_by(exp) %>%
  summarise(
    n=n(),
    age_years = sprintf('%.1f (%.1f)', mean(age), sd(age)),
    males = sprintf('%i (%.1f %%)', sum(gender=='male'), 100*sum(gender=='male')/n())
  ) #%>%
  # kable() %>% 
  # kable_styling(bootstrap_options = "striped", full_width = F)
```

Supplementary per-group info stratified by level and stimType descriptives:
```{r}
x = D %>%  
  # One row per participant
  group_by(id) %>%
  slice(1) %>%
  
  # Per stimType-level combination
  group_by(stimType, level) %>%
  summarise(
    n=n(),
    males = sprintf('%i (%.1f %%)', sum(gender=='male'), 100*sum(gender=='male')/n()),
    age_years = sprintf('%.1f (%.1f)', mean(age), sd(age))
  )

y = D %>%  # All trials
  group_by(stimType, level) %>%
  summarise(
    arithmetic = sprintf('%.1f%%', mean(equationCorrectness)*100),
    recall = sprintf('%.1f%%', mean(recallProportion)*100)
  )

bind_cols(x, y[,3:4])# %>%
  # kable() %>% 
  # kable_styling(bootstrap_options = "striped", full_width = F)
```


# Concurrent load: Figure 2
```{r figure2, fig.height=5, fig.width=4}
# Fit a simple mixed model to show the results while subtracting individual differences
fit_full = lmer(pggInvest ~ span*stimType + (1|id), D)

# Add the fits to the data
x = ranef(fit_full)  # random effects for each subject
D$pggInvestOffset = mapvalues2(D$id, from=rownames(x$id), to=x$id$`(Intercept)`)  # map it unto data
df_tmp = filter(D, span==1 & level=='1-7')
y_offset = mean(df_tmp$pggInvest - df_tmp$pggInvestOffset, na.rm=T)
D$pggInvestZeroCenter = D$pggInvest - D$pggInvestOffset - y_offset

# Plot the figure
figure2 = ggplot(D, aes(x=span, y=pggInvestZeroCenter, color=level)) + 
  stat_summary(fun.data='mean_cl_boot', position=position_dodge(0.5), size=0.7) +
  stat_summary(fun.y=mean, geom="line", position=position_dodge(0.5), lwd=1.3) +
  #facet_grid(~stimType) +
  labs(title='WM load and cooperation', y='Relative Public Goods Game Investment', x='Complex Span') + 
  scale_x_continuous(breaks=1:7) + scale_y_continuous(breaks=seq(-100, 100, 2))

figure2 = style_my_plot(figure2)

# Show it and save it
ggsave('figures/Figure 2 - PGG and CS span.png', figure2, width=6, height=6, units='cm', dpi=300, scale=1.7)
figure2
```

Supplementary figure:
```{r}
figureS2 = ggplot(D, aes(x=span, y=pggInvest, color=level)) + 
  stat_summary(fun.data='mean_cl_boot', position=position_dodge(0.5), size=0.3) +
  stat_summary(fun.y=mean, geom="line", position=position_dodge(0.5)) +
  facet_grid(~stimType) +
  labs(title='WM load and cooperation', y='Public Goods Game Investment', x='Complex Span') + 
  scale_x_continuous(breaks=1:7) + scale_y_continuous(breaks=seq(-100, 100, 2))

figureS2 = style_my_plot(figureS2)

# Show it and save it
ggsave('figures/Figure S2 - PGG and CS span.png', figureS2, width=9, height=6, units='cm', dpi=300, scale=1.7)
figureS2
```




# Concurrent load: inference
Main test. Output of `LRT` consists of:

1. LRT test to obtain p-value
2. Parameters from `lmer` of the full model
3. Bootstrapped parameter estimates and intervals
4. BIC-based Bayes Factor based on the BIC values from (1)

```{r pgg load, message=FALSE, warning=FALSE}
LRT(D, 
    pggInvest ~ span + stimType + time_hours + (span|id), 
    pggInvest ~    1 + stimType + time_hours + (span|id))
```

I would looove to go proper Bayesian, but it is too slow:
```{r}
# # Bayesian version (takes a looooooong time to run!)
# # Needs well-considered priors
# library(brms)
# full = brm(pggInvest ~ span + stimType + time_hours + (span|id), D, chains=1, cores=1, iter=650, warmup=150, save_all_pars=TRUE)
# null = brm(pggInvest ~    1 + stimType + time_hours + (span|id), D, chains=1, cores=1, iter=650, warmup=150, save_all_pars=TRUE)
# bayes_factor(full, null)
# loo(full, null)


# # Disregard the many responses at 0 and 100 as being from a different process
# D$pggInvest_scaled = D$pggInvest / 100  # Scale down to [0, 1]; effectively making this equivalent to a stretched beta
# 
# Q = bf(
#   pggInvest_scaled ~ span + (1|id),
#   phi ~ span,
#   zoi ~ span,
#   coi ~ span,
#   family = zero_one_inflated_beta()
# )
# 
# Q_fit = brm(Q, D)
```


Effect of CS stimulus type - does it change (interact with) the slope?
```{r}
LRT(D, 
    pggInvest ~ span * stimType + time_hours + (span|id), 
    pggInvest ~ span + stimType + time_hours + (span|id))
```

Effect of difficulty level - does it change (interact with) the slope?
```{r}
LRT(D, 
    pggInvest ~ span * level + stimType + time_hours + (span|id),
    pggInvest ~ span + level + stimType + time_hours + (span|id))
```




# Figure 3: Effect of time
```{r figure3}
# Get subject random effects. Intercept t=0, but keep level-specific offsets
fit = lmer(pggInvest ~ time_secs + level + stimType + (1|id), D)
x = ranef(fit)

# map it unto data
D$pggInvestOffset = mapvalues2(D$id, from=rownames(x$id), to=x$id$`(Intercept)`)

# Plot linearly with data
figure3 = ggplot(D, aes(x=time_secs/60, y=pggInvest - pggInvestOffset, color=level, fill=level)) +  # plot data with subject-specific offsets
  stat_smooth(method='glm') +  # Make it linear
  stat_summary_bin(fun.y=mean, geom='point', binwidth=2, size=2) +
  facet_grid(stimType~level) + 
  
  # Styling
  labs(title='Load over time and cooperation', y='Public Goods Game Investment', x='Minutes elapsed') + 
  scale_x_continuous(breaks=seq(-10, 30, 5)) + coord_cartesian(xlim=c(0, 17), ylim=c(35, 60)) + scale_y_continuous(breaks=seq(-100, 100, 5))

figure3 = style_my_plot(figure3)
figure3

ggsave('figures/Figure 3 - PGG and CS over time.png', figure3, width=8, height=4.5, units='cm', dpi=300, scale=2.5)
```


# Depletion on PGG investment: Inference
Main analysis:

```{r}
LRT(D, pggInvest ~ level * time_hours + stimType + (span|id),
       pggInvest ~ level + time_hours + stimType + (span|id))
```

Without easy-letters:
```{r}
LRT(filter(D, !(level == '1-3' & stimType=='Letters')), 
    pggInvest ~ level * time_hours + stimType + (span|id),
    pggInvest ~ level + time_hours + stimType + (span|id))
```



Only easy-letters
```{r}
LRT(filter(D, (level == '1-3' & stimType=='Letters')), 
    pggInvest ~ time_hours + (span|id),
    pggInvest ~         1 + (span|id))
```


# Supplementary stuff
See the notebook for the dots task for more supplementary analyses.


